---
title: "Data Sci: Group 10 Draft/Outline"
author: "Matthew O'Donnell, Sean Sander, Daniel Forcade"
date: "4/8/2021"
output: pdf_document
---

```{r setup, include=FALSE, echo = FALSE, warning = FALSE}
library(rmdformats)
```
```{r, warning = FALSE, echo = FALSE, message = FALSE}
library("ggplot2")
library("dplyr")
library("gridExtra")
library("quantreg")
library("gganimate")
library("animation")
library("transformr")
library("viridis")
library("tidyverse")
library("ggsci")
library("ggpubr")
library("gganimate")
library("magick")
library("ggthemes")
library("tidyverse")
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_chunk$set(fig.width=10, fig.height=10)  
```


## 1.  Introduction:

* What is the source of the data?: *Answer*
* Where and when was it created?: *Answer*
* Is it a sample: *No*
* Do you suspect any sampling bias: *Answer*
* Was it an experimental or an observational study: *Observational - not a study*
* How were measurements taken: *NYPD Data*
* Do you suspect any bias in the questions or measurements: *The exploration of this question is an auxiliary part of our analysis*
* Why is this data of interest to you: *Answer*
* What kind of data cleaning was necessary: *Answer - found in data cleaning section*





## 2.  Data Visualizations:  















## 3.  Machine Learning: To Do Later:

Ideas:

* Compare precincts against normal distribution
* Compare precincts against eachother

* Compare officer race vs complaint outcome
* Compare complainant race vs complaint outcome













## 4.  Conclusions:  







## 5.  Limitations/Recommendations:  

* NYPD had a large amount of time to manipulate data before public release
* Second organization accessed data-set before public release (increased potential for manipulation)

* Research into precinct locations would allow borough or location partition for data analysis






## Names of Variables in Data Set:
```{r}
allegations <- read.csv("allegations_202007271729.csv")

names(allegations)
```


## New






































## 1.  
Ran a histogram of complaint type, separated by year.  Noticed that years 1985 ~ 2000 have a limited amount of data.  May need to research as to why.

```{r, warning = FALSE, message = FALSE}

### Complaint Type Histogram
ggplot(data = allegations,
       mapping = aes(x = fado_type, fill = fado_type, alpha = .8))+
  geom_histogram(stat = "count")+
  facet_wrap(~ year_received)+
  labs(title = "Complaint Type Histogram through Year", x = "Complaint Type", y = "Count", caption = "Figure 1", subtitle = "1985-2020")+
  theme_dark()+
  theme(axis.text.x = element_text(angle = 90, vjust = -.1))+
  scale_fill_tron(name = "Complaint Type")+
  guides(alpha = FALSE)
```

## 2.  
Figure 2 is a density bar graph of complaint ethnicity, separated by complaint type.  I wanted to see if there were any trends in complaints depending on ethnicity of the victim.  
I noticed in this graph that there are a lot of _____ or 'blank', 'unfilled' categories in this data set.  These are in addition to several 'NA' or 'Unknown' categories.  We will likely have to decide on a path on how to rewrite, categorize, or otherwise parse the missing data.

```{r, warning = FALSE, message = FALSE}
## complaint ethnicity
ggplot(data = allegations,
       mapping = aes(x = complainant_ethnicity, fill = complainant_ethnicity,
                     alpha = .9))+
  geom_bar(color = "black")+
  facet_wrap(~fado_type)+
  labs(title = "Complaintant Ethnicity Bar Chart", subtitle = "Seperated by Complaint Type.  Note: First Value is Blank", caption = "Figure 2", x = "Complaint Ethnicity")+
  theme_dark()+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_discrete(name = "Complaint Ethnicity")+
  guides(alpha = FALSE)
```

## 3.  
Figure 3 is a density bar graph of `board_disposition`, which I believe to be the outcome of the complaint - separated by `mos_ethnicity`, which I believe to be officer ethnicity.  
A quick research into the data lead me to believe that 'mos' in this data set stands for **Member of Service**, or police officer.  
Here I was checking for patterns if there were any obvious outcome bias depending on officer ethnicity.

```{r, warning = FALSE, message = FALSE}
## Officer Race by Board Disposition
ggplot(data = allegations,
       mapping = aes(x = board_disposition, fill = board_disposition))+
  geom_bar(color = "black")+
  facet_wrap(~mos_ethnicity)+
  theme_cleveland()+
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  labs(title = "Officer Complaint Outcome Bar Chart", subtitle = "Seperated by Officer Ethnicity", caption = "Figure 3", x = "Board Disposition (Complaint Outcome)")+
  scale_fill_futurama(name = "Complaint Outcome")
```
  
  
## 4.  
Here I did a quick scan on compliant type by `complainant_gender`.  As the graph-set shows, there are a lot of different and unknown gender variables for complainants.  If we want to work with complainant gender, we may want to limit or parse the 4 low-value and/or missing categories (Transman, Not described, Transwoman, Gender non-conforming)

```{r, warning = FALSE, message = FALSE}
## Gender Complaints

ggplot(data = allegations,
       mapping = aes(x = fado_type, fill = fado_type,
                     alpha = .9))+
  geom_bar(color = "black")+
  facet_wrap(~complainant_gender)+
  labs(title = "Complaint Type", subtitle = "Seperated by Gender", caption = "Figure 4", x = "Complaint Type")+
  theme_gdocs()+
  theme(axis.text.x = element_text(angle = 90))+
  scale_fill_jco(name = "Complaint Type")+
  guides(alpha = FALSE)
```